---
title: "titanic survial"
author: "Biswajit Maharathi"
date: "April 10, 2016"
output: html_document
---

This scripts creates a random forest model to fit the titanic data given in kaggle competition and then predicts the survival rate for the testing data set.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r,warning=FALSE,results='hide'}
## load libraries 
library(ggplot2); library(randomForest); library(caret);library(dplyr);library(mice)
```

Load the required datasets 
```{r}
##load datasets
training <-read.csv("train.csv", header=TRUE, sep = ",")
testing  <-read.csv("test.csv", header=TRUE, sep = ",")
```
Merge both data sets training and testing for data cleaning 
```{r}
##combine data sets 
totaldata <- bind_rows(training,testing)
```
```{r}
names(totaldata)
##[1] "PassengerId" "Survived"    "Pclass"      "Name"        "Sex"         "Age"         "SibSp"      
##[8] "Parch"       "Ticket"      "Fare"        "Cabin"       "Embarked"   
```

```{r}
## fields with NA value : Age(263) ; Fare (1) ;Embarked 
```
Replace the missing fare value with median fare of the dataset.
```{r}
totaldata[is.na(totaldata$Fare),]$Fare <- median(totaldata$Fare,na.rm=TRUE)
totaldata[totaldata$Embarked=='',]$Embarked <- 'C'
```
Insert age for the missing age values. 
```{r}
#age correction 
ft_var <- c('PassengerId','Pclass','Sex','Embarked')
totaldata[ft_var] <- lapply(totaldata[ft_var], function(x) as.factor(x))

set.seed(234)
mice_mod <- mice(totaldata[, !names(totaldata) %in% c('PassengerId','Name','Ticket','Cabin','Survived')], method='rf') 
mice_output <- complete(mice_mod)
#replace the age
totaldata$Age <- mice_output$Age
```
Now split the data to training and testing for modelling 

```{r}
##split data again to train and test
trainfinal <- totaldata[1:891,]
testfinal <- totaldata[892:1309,]
```

Build the predictive model(using random forest)
```{r}
##build prediction model
set.seed(556)
rf_model <- randomForest(factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch + 
                             Fare + Embarked,data = trainfinal)
##model error 
plot(rf_model, ylim=c(0,0.36))
legend('topright', colnames(rf_model$err.rate), col=1:3, fill=1:3)

```
Predict the values for training set and check the confusion matrix 
```{r}
##predict 
prd <- predict(rf_model,trainfinal)
confusionMatrix(prd,trainfinal$Survived)
```

Predict the required values in the testing data set. 
```{r}
##predict the final required values 
prdfinal <- predict(rf_model,testfinal)
```
Write the data to a file 
```{r}
#write to the file 
#submit <- data.frame(PassengerId = testfinal$PassengerId)
#submit$Survived <- prdfinal

#write.csv(submit, file = "titanic_survive_rf_v02.csv", row.names=FALSE)
```


